WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2015-07-08 ms.tex
Add area histogram

The \ac{BAYESTAR} adaptive sampling process works as follows. W begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 3072$ pixels of an $N_{\mathrm{side},0}=16$ \ac{HEALPix} grid. At this resolution, each pixel has an area of 13.4~deg$^2$. We then rank the pixels by contained probability (assuming constant probability density within a pixel), and subdivide the most probable $N_{\mathrm{pix},0}/4$ pixels into $N_{\mathrm{pix},0}$ new daughter pixels. We then evaluate the posterior again at the centers of the new daughter pixels, sort again, and repeat seven times. After the last iteration, we have evaluated the posterior probability density a total of $8 N_{\mathrm{pix},0}$ times. On \emph{most} subdivision steps, we descend one level deeper in \ac{HEALPix} resolution. This process is illustrated in Figure~\ref{fig:adaptive-illustration}.

@@

Within each iteration, all of the marginalization integrals may be evaluated in parallel. In our C language implementation, the inner loop over pixels is parallelized with OpenMP\footnote{http://openmp.org/}.

@@

\subsection{Self-consistency of credible levels}

One might naively expect that self\nobreakdashes-consistency would require the 90\% confidence area and searched area histograms to intersect at 90\% of detections, but this is not generally required because the posteriors of different events have widely different dimensions. However, it is true that 90\% of sources should be found within their respective 90\% confidence contours. This can be formalized into a graphical self\nobreakdashes-consistency test called a probability\nobreakdashes--probability ($P$\nobreakdashes--$P$) plot (see \citealt{SiderySkyLocalizationComparison} for applications in \ac{GW} parameter estimation). For each event, one follows the water\nobreakdashes-filling procedure to walk outward from the most probable pixel until one reaches the pixel that contains the true sky location, noting the total probability $p$ that has been traversed.

>>>>

The \ac{BAYESTAR} adaptive sampling process works as follows. We begin by evaluating the posterior probability density at the center of each of the $N_{\mathrm{pix},0} = 3072$ pixels of an $N_{\mathrm{side},0}=16$ \ac{HEALPix} grid. At this resolution, each pixel has an area of 13.4~deg$^2$. We then rank the pixels by contained probability (assuming constant probability density within a pixel), and subdivide the most probable $N_{\mathrm{pix},0}/4$ pixels into $N_{\mathrm{pix},0}$ new daughter pixels. We then evaluate the posterior again at the centers of the new daughter pixels, sort again, and repeat seven times. After the last iteration, we have evaluated the posterior probability density a total of $8 N_{\mathrm{pix},0}$ times. On \emph{most} subdivision steps, we descend one level deeper in \ac{HEALPix} resolution. This process is illustrated in Figure~\ref{fig:adaptive-illustration}.

@@

Within each iteration, all of the marginalization integrals may be evaluated in parallel. In our C language implementation of \ac{BAYESTAR}, the inner loop over pixels is parallelized with OpenMP\footnote{http://openmp.org/}.

@@

All simulated sources have component masses distributed uniformly between 1.2 and 1.6\,$M_\odot$ and randomly oriented spins with dimensionless magnitudes $\chi$ between -0.05 and +0.05. Sky positions and binary orientations are random and isotropic. Distances are drawn uniformly from ${D_\mathrm{L}}^3$, reflecting a uniform source population (neglecting cosmological effects, which are small within the Advanced \acs{LIGO} \ac{BNS} range).

The simulated waveforms were deposited in Gaussian noise that has been filtered to have the \acp{PSD} consistent with \cite{LIGOObservingScenarios}. They were detected using the real\nobreakdashes-time matched\nobreakdashes-filter pipeline, GSTLAL\_INSPIRAL \citep{Cannon:2011vi}. Candidates with estimated \acp{FAR} less than $10^{-2}$\,yr$^{-1}$ were considered to be detections. Because using Gaussian noise results in lower \acp{FAR} than would be calculated in realistically glitchy detector noise, we imposed an additional detection threshold on the network \ac{SNR}, $\hat{\rho} \geq 12$, which has been found to correspond to a comparable \ac{FAR} in the initial \ac{LIGO} runs. For further analysis of the effect of glitchy noise on detection and sky localization, see \cite{BerryLocalization}.

\subsection{Areas}

\begin{figure*}
    \includegraphics{area-hist}
    \caption{\label{fig:area-hist}Cumulative histograms of sky area, broken down by observing run and detector network. The plots in the left column show the 90\% credible area and the plots in the right column show the searched area. From top bottom, the rows refer to the following observing scenarios/network configurations: 2015/HL, 2016/HL, 2016/HV, 2016/LV, 2016/HLV. The shaded regions represent the 95\% confidence bounds. The magenta lines are represent \ac{BAYESTAR} and the blue lines LALINFERENCE. Where relevant, dotted lines show all events in the given network configuratoin and solid lines show only events for which the matched\nobreakdashes-filter pipeline triggered on all operating detectors. Note that statistically significant differences in areas between the \ac{BAYESTAR} and LALINFERENCE localizations occur only for events that were below the detection threshold in one or more detectors.}
\end{figure*}

\subsection{Self-consistency}

\textbf{need to rewrite this paragraph, copied from first2years paper} One might naively expect that self\nobreakdashes-consistency would require the 90\% confidence area and searched area histograms to intersect at 90\% of detections, but this is not generally required because the posteriors of different events have widely different dimensions. However, it is true that 90\% of sources should be found within their respective 90\% confidence contours. This can be formalized into a graphical self\nobreakdashes-consistency test called a probability\nobreakdashes--probability ($P$\nobreakdashes--$P$) plot (see \citealt{SiderySkyLocalizationComparison} for applications in \ac{GW} parameter estimation). For each event, one follows the water\nobreakdashes-filling procedure to walk outward from the most probable pixel until one reaches the pixel that contains the true sky location, noting the total probability $p$ that has been traversed.
